---
title: "Codebook"
author: "Rene Marves"
date: "26 de febrero de 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Codebook

Codebook was generated on `r as.character(Sys.time())` during the same process that generated the dataset. See README.md or README.html for details on dataset creation.

## Variable list and descriptions

|Variable name   |Description                                          |
|----------------|-----------------------------------------------------|
|Subject         |ID the subjectwho performed the activity             |
|                |for each window sample. Its range is from 1 to 30.   |
|ActivityName    |Activity name                                        |


The rest of the variables follow the format:

(Measurement Dimension)(Mesurement subject)(Instrument)(Measurement Type)

Where:

|Item                   |Description                                   |
|-----------------------|----------------------------------------------|
|Measurement Dimension  |Can be: frequency or time                     |
|Measurement subject    |Can be: Body, Gravity or BodyBody             |
|Instrument             |Can be: Acceleromenter o Gyroscope            |
|Measurment type        |Can be: Mean or Standard Deviation of a X, Y, |
                        |or Z component, or a magnitude; or a jerk     |
                        |measurement therof                            |

Table: All other variables

## Dataset structure

```{r}
str(dtTidy2)
```

## Show a few rows of the dataset

```{r}
dtTidy2
```

## Summary of variables

```{r}
summary(dtTidy2)
```

## Save to file

Save data table objects to 2 tab-delimited text files called 
**`r filename1`** and 
**`r filename2`**
